\thebibliography{99}
\bibitem{Liu2012}
Bing Liu,
\textit{Sentiment Analysis and Opinion Mining},
Morgan \& Claypool Publishers, 2012.

\bibitem{TwitterStats}
InternetLiveStats.com,
\textit{Twitter Usage Statistics},
\url{https://www.internetlivestats.com/twitter-statistics}, dostęp: maj 2025.

\bibitem{Sarcasm}
Yiyi Liu, Yequan Wang, Aixin Sun, Xuying Meng, Jing Li, Jiafeng Guo,
\textit{A Dual-Channel Framework for Sarcasm Recognition by Detecting Sentiment Conflict},
arXiv:2109.03587, 2021.

\bibitem{Emotions}
Nandwani, P., Verma, R,
\textit{A review on sentiment analysis and emotion detection from text. Soc. Netw. Anal. Min. 11, 81},
\url{https://doi.org/10.1007/s13278-021-00776-6}, 2021.

\bibitem{NLPHistory1}
Dan Jurafsky and James H. Martin,
\textit{Speech and Language Processing (3rd ed. draft)},
\url{https://web.stanford.edu/~jurafsky/slp3}, 2025.

\bibitem{NLPHistory2}
Johri, Prashant and Khatri, Sunil Kumar and Al-Taani, Ahmad and Sabharwal, Munish and Suvanov, Shakhzod and Chauhan, Avneesh,
\textit{Natural Language Processing: History, Evolution, Application, and Future Work},
\url{https://doi.org/10.1007/978-981-15-9712-1_31}, 2021.

\bibitem{NLP2017}
Diksha Khurana, Aditya Koli, Kiran Khatter, Sukhdev Singh,
\textit{Natural Language Processing: State of The Art, Current Trends and Challenges},
arXiv:1708.05148, 2017.

\bibitem{SentA2020}
Sharat Sachin, Abha Tripathi, Navya Mahajan, Shivani Aggarwal, Preeti Nagrath,
\textit{Sentiment Analysis Using Gated Recurrent Neural Networks},
\url{https://doi.org/10.1007/s42979-020-0076-y}, 2020.

\bibitem{BERT}
Jacob Devlin, Ming-Wei Chang, Kenton Lee, Kristina Toutanova,
\textit{BERT: Pre-training of Deep Bidirectional Transformers for Language Understanding},
arXiv:1810.04805, 2018.

\bibitem{RoBERTa}
Yinhan Liu, Myle Ott, Naman Goyal, Jingfei Du, Mandar Joshi, Danqi Chen, Omer Levy, Mike Lewis, Luke Zettlemoyer, Veselin Stoyanov,
\textit{RoBERTa: A Robustly Optimized BERT Pretraining Approach},
arXiv:1907.11692, 2019.

\bibitem{ALBERT}
Zhenzhong Lan, Mingda Chen, Sebastian Goodman, Kevin Gimpel, Piyush Sharma, Radu Soricut,
\textit{ALBERT: A Lite BERT for Self-supervised Learning of Language Representations},
 arXiv:1909.11942, 2020.

\bibitem{DistilBERT}
Victor Sanh, Lysandre Debut, Julien Chaumond, Thomas Wolf,
\textit{DistilBERT, a distilled version of BERT: smaller, faster, cheaper and lighter},
arXiv:1910.01108, 2020.

\bibitem{OracleBayes}
International Business Machines Corporation,
\textit{Oracle Naive Bayes},
\url{https://www.ibm.com/docs/pl/spss-modeler/saas?topic=mining-oracle-naive-bayes}, dostęp: marzec 2025.

\bibitem{TextClassOnBigData}
D.C. Asogwa, S.O. Anigbogu, I.E. Onyenwe, F.A. Sani,
\textit{Text Classification Using Hybrid Machine Learning Algorithms on Big Data},
arXiv:2103.16624, 2021.

\bibitem{HateClassification}
D.C Asogwa, C.I Chukwuneke, C.C Ngene, G.N Anigbogu,
\textit{Hate Speech Classification Using SVM and Naive BAYES},
arXiv:2204.07057, 2022.

\bibitem{PLMsvsSVM}
Yasmen Wahba, Nazim Madhavji, John Steinbacher,
\textit{A Comparison of SVM against Pre-trained Language Models (PLMs) for Text Classification Tasks},
arXiv:2211.02563, 2022.

\bibitem{SAKapur}
Keshav Kapur, Rajitha Harikrishnan,
\textit{Comparative Study of Sentiment Analysis for Multi-Sourced Social Media Platforms},
arXiv:2212.04688, 2022.

\bibitem{SCMohamed}
Mohamed Kayed, Rebeca P. Díaz-Redondo, Alhassan Mabrouk,
\textit{Deep Learning-based Sentiment Classification: A Comparative Survey},
arXiv:2312.17253, 2023.

\bibitem{EmotionTra}
Mahdi Rezapour,
\textit{Emotion Detection with Transformers: A Comparative Study},
arXiv:2403.15454, 2024.

\bibitem{LSTMandBERT}
Pankaj M Thakur and Tejas M,
\textit{Multi-class Classification of Twitter Sentiments using Frequency based, LSTM and BERT Methods},
\url{https://api.semanticscholar.org/CorpusID:248629612}, 2022.

\bibitem{TTSD}
Kaggle Datasets,
\textit{Twitter Tweets Sentiment Dataset},
\url{https://www.kaggle.com/datasets/yasserh/twitter-tweets-sentiment-dataset}, licencja: CC0 Public Domain, dostęp: listopad 2024.

\bibitem{MSAD}
Shahriar Parvez,  
\textit{Multiclass Sentiment Analysis Dataset},  
\url{https://huggingface.co/datasets/Sp1786/multiclass-sentiment-analysis-dataset}, licencja: Apache 2.0, dostęp: listopad 2024.

\bibitem{TweetEval}
Sara Rosenthal, Noura Farra, Preslav Nakov,
\textit{SemEval-2017 Task 4: Sentiment Analysis in Twitter},
Proceedings of the 11th International Workshop on Semantic Evaluation (SemEval-2017), pp. 502--518, 2017.
\url{https://huggingface.co/datasets/cardiffnlp/tweet_eval}, licencja: CC BY 3.0, dostęp: listopad 2024.

\bibitem{EmotionDs}
Marcos Penelas,  
\textit{Emotion Dataset},  
\url{https://www.kaggle.com/datasets/marcospenelas/emotion}, licencja: MIT, dostęp: listopad 2024.

\bibitem{PreprocessingKurniasih}
Aliyah Kurniasih and Lindung Parningotan Manik,
\textit{On the Role of Text Preprocessing in BERT Embedding-based DNNs for Classifying Informal Texts},
International Journal of Advanced Computer Science and Applications(IJACSA), 13(6)
\url{http://dx.doi.org/10.14569/IJACSA.2022.01306109}, 2022.

\bibitem{PreprocessingTan}
Tan, Kian \& Lee, Chin-Poo \& Lim, Kian,
\textit{RoBERTa-GRU: A Hybrid Deep Learning Model for Enhanced Sentiment Analysis},
Applied Sciences. 13. 3915. 10.3390/app13063915, 2023.

\bibitem{ClassImbalance}
Sophie Henning, William Beluch, Alexander Fraser, Annemarie Friedrich,
\textit{A Survey of Methods for Addressing Class Imbalance in Deep-Learning Based Natural Language Processing},
arXiv:2210.04675, 2022.

\bibitem{UczMasz}
Aurélien Géron,
\textit{Uczenie maszynowe z użyciem Scikit-Learn, Keras i TensorFlow},
Helion, 2023.

\bibitem{SKLearnSVM}
Scikit-learn developers,
\textit{SVC - Support Vector Classification},
\url{https://scikit-learn.org/stable/modules/generated/sklearn.svm.SVC.html}, dostęp: luty 2025.

\bibitem{KernelComparison}
Kalcheva, Neli and Karova, Milena and Penev, Ivaylo,
\textit{Comparison of the accuracy of SVM kernel functions in text classification},
\url{https://doi.org/10.1109/BIA50171.2020.9244278}, 2020.

\bibitem{LSTMIntroduction}
Hochreiter, Sepp and Schmidhuber, Jürgen,
\textit{Long Short-Term Memory},
\url{https://doi.org/10.1162/neco.1997.9.8.1735}, 1997.

\bibitem{PyTorchDocs}
PyTorch Contributors,
\textit{PyTorch documentation},
\url{https://pytorch.org/docs/stable/index.html}, dostęp: luty 2025.

\bibitem{KerasDocs}
Keras Contributors,
\textit{Keras Documentation},
\url{https://keras.io/api/}, dostęp: luty 2025.

\bibitem{GRUvsLSTM}
Junyoung Chung, Caglar Gulcehre, KyungHyun Cho, Yoshua Bengio,
\textit{Empirical Evaluation of Gated Recurrent Neural Networks on Sequence Modeling},
arXiv:1412.3555, 2014.

\bibitem{HuggingFaceDocs}
Hugging Face,
\textit{Transformers Documentation},
\url{https://huggingface.co/docs/transformers/en/index}, dostęp: styczeń 2025.

\bibitem{Bias1}
Tolga Bolukbasi, Kai-Wei Chang, James Zou, Venkatesh Saligrama, Adam Kalai,
\textit{Man is to Computer Programmer as Woman is to Homemaker? Debiasing Word Embeddings},
arXiv:1607.06520, 2016.

\bibitem{Bias2}
Ninareh Mehrabi, Fred Morstatter, Nripsuta Saxena, Kristina Lerman, Aram Galstyan,
\textit{A Survey on Bias and Fairness in Machine Learning},
arXiv:1908.09635, 2022.

\bibitem{Bias3}
Tony Sun, Andrew Gaut, Shirlyn Tang, Yuxin Huang, Mai ElSherief, Jieyu Zhao, Diba Mirza, Elizabeth Belding, Kai-Wei Chang, William Yang Wang,
\textit{Mitigating Gender Bias in Natural Language Processing: Literature Review},
arXiv:1906.08976, 2019.

\endinput